---
title: "Basic Operations in R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##### introduction to R########
### creating Vectors/Variables####

X <- 10:20 # 10 to 20.
```{r}
X <-  10:20 # 10 to 20. #it is used to assign vectors to X just like = in python.
```
```{r}
X
```
```{r}
y <- c(1,2,3,4,5,6,7,8,9,10,11) # c is a function and it means combine
```
```{r}
y
```
####### arithmetic operations #######
```{r}
X+y #addition
```
```{r}
y=X+2 # updating y
```
```{r}
y
```
```{r}
z <- y*3
```

```{r}
z
```

```{r}
subtotal=z-6
```
```{r}
subtotal
```
```{r}
answer <- subtotal/3
```
```{r}
answer
```
```{r}
X
```
```{r}
answer=((X+2)*3-6)/3
```
```{r}
answer
```
```{r}
x=10
```
```{r}
lab=((x+2)*3-6)/3
```
```{r}
lab
```
### Data Types and assigning data to a vector/variable#####
```{r}
cards=c('tiffany','funlola','yemi','makwin','fatimat') # c is a function to concatenate
```

```{r}
cards
```
```{r}
typeof(cards) # to check for the data type.
```
```{r}
number <- c(3000,NA,4000,3000,5000)
```
```{r}
typeof(number)
```
```{r}
play=c(cards,number)
play
```
```{r}
typeof(play)
```
```{r}
monster=c(T,T,T,F,T,T,T,T,T)
```
```{r}
typeof(monster)
```
```{r}
print(monster)
```
```{r}
number=c(number,monster)
```
```{r}
number
```
```{r}
typeof(number)
```
```{r}
print(number)
```
```{r}
coerce.check <- c(play,monster)
coerce.check
```
```{r}
typeof(coerce.check)
```
```{r}
y <- -7
```

```{r}
rm(y) #to remove a given variable from the Global eniroment
```

```{r}
ls()  #to check for active variables in the Global enviroment
```
```{r}
number <-c(1,7,6,4,9,10,12,23)
print(number)
```

```{r}
intern_name <-c("tiffany","funlola","yemi","makwin","fatmimat")
print(intern_name)
```

```{r}
num_vec_1 <- c(1, 4, 9)
num_vec_2 <- c("1","4","9")
```

```{r}
typeof(num_vec_1)
typeof(num_vec_2)
```
```{r}
scores <- c(3000, 4000, 5000, 6000, 7000)
names(scores) # used to check for column names
names(scores) <- intern_name # making the intern name the column names
attributes(scores)
names(scores)
#names and attributes carry out the same function.but like info and dtypes in python.attributes returns more than column names 
scores
```
```{r}
intern_name <- c("tiffany" = 3000, "funlola"=4000, "yemi" = 5000, "makwin" = 6000, "fatmimat" = 7000)
print(intern_name)
```

Another Method 
```{r}
age <- c(21, 26, 32, 31, 28)
attributes(age)
```
```{r}
names(age) <- c("tiffany","funlola","yemi","makwin","fatmimat")
age
```
#Descriptive Statistics

```{r}
number
sum(number) # sum of all the numbers
```
```{r}
mean(number
     )
```
```{r}
median(number)
```

```{r}
length(number) # no of entries
```
```{r}
var(number)
```
```{r}
sd(number)
```
```{r}
round(sd(number)) # rounding up the sd
```
```{r}
class_score <- c("Clara" = 97, "Solomon" = 100, "Alex" = 67, "Dr Tito" = 78, "Duchima" = 76, "Steve" = 78, "Stepen" = 89, "Dr T" = 87, "Chinwe" = 78, "Victor" = 79, "Patricia" = 79, "Precious" = 84)
print(class_score)
```

```{r}
mean(class_score)
median(class_score)
sd(class_score)
var(class_score)
```

```{r}
args(sample) #display the aurgement of a fuction i.e other parameters that can be used with the fuction
```
```{r}
args(mean)
```
```{r}
?sample # ? does the same thing as shift tab in python

```


```{r}
?mean
```

```{r}
x <- c(0:10, 50)
xm <- mean(x)
c(xm, mean(x, trim = 0.10))
```

```{r}
args(median)

```
```{r}
args(median)
```

```{r}
number <- c(3000, NA, 4000, 3000, 5000)
number
```
```{r}
mean(number) 
median(number)
```

```{r}
mean(number, na.rm = TRUE) # na.rm means without missing values
median(number, na.rm = TRUE)
```
# Slicing and Indexing a Vector in R
```{r}
age <- c(22, 25, 28, 29, 21)
```


```{r}
name<- c("tiffany","funlola","yemi","makwin","fatmimat")
```

```{r}
name[4] # 4th entry
```
```{r}
name[-1] # return all entries except the 1st one
```

```{r}
name[c(1,3,5)]

```

```{r}
name[c(-1,-3)]
```
```{r}
name[-c(1,3)]
```
```{r}
name
```
```{r}
names(age) <- name
age
```

```{r}
age["tiffany"]
```

```{r}
age[c("fatmimat", "yemi")]
```
```{r}
age[1:3]
```
```{r}
age[c(1,2,3)]
```
#Creating a vector using 'seq'
```{r}
weight <- seq(10, 100, by = 10) # sequence of 10-100 but difference of 10
weight
```
```{r}
weight[-(4:6)]
```

```{r}
weight[weight>70]
```

```{r}
weight[weight>50]
```
#Working with Matrices in R
```{r}
weight <- seq(10, 120, by = 10)
dim(weight)
dim(weight) <- c(3,4) # the dimension just like reshape
weight
```

```{r}
mtr <- matrix(1:12, nrow = 3)
mtr
```

```{r}
mtr <- matrix(1:12, ncol = 4)
mtr
```

```{r}
mtr <- matrix(1:12, ncol = 4, byrow = T) #it means arrange it rowwise
mtr
```

```{r}
math <- c(61, 53, 34, 45, 55)
ict <- c(73, 58, 67, 59, 86)
```

```{r}
subject <- cbind(math,ict) #binding by column i.e concat the ict comes to meet the maths
subject
```
```{r}
rownames(subject) <- c("2015", "2016", "2017", "2018", "2019") # function for assiging row names
subject
```

```{r}
subject <- t(subject) #Transpose i.e make column the rows and rows the column
subject
```

```{r}
english <- c(76, 56, 87, 95, 67) 
```

```{r}
subject <- rbind(subject, english) # rbind means rowbind
subject
```
# Creating a Matrix with a single line of code
```{r}
weight <- matrix(c(66.5, 56.0, 75.0, 65.4, 67.9, 70.9, 81.2, 50.0, 76.3 ),
        nrow = 3, byrow = T,
        dimnames = list(c("tiffany", "makwin",  "fatmimat"),
                        c("2019", "2020", "2021")))

weight
```
```{r}
data <- matrix(1:10, nrow = 4, ncol = 4)
data
```

```{r}
Gideon <- matrix(c(2:10), nrow = 4, ncol = 4, byrow = T,
                      dimnames = list(c("row_1", "row_2", "row_3", "row_4"),
                                     c("col_1", "col_2", "col_3", "col_4")))
Gideon                     
```
### indexing elements in matrix
# matrices are natural extensions of vectors.
# matrices are multidimensional vectors.

```{r}
unit <- c(231,323,450,500,600,700,654,321,170)
sales=matrix(unit,nrow=3,byrow=T)
sales
```
```{r}
sales[1,3]
```
```{r}
sales[1] # single indexing
```
```{r}
sales[1,] # row indexing
```
```{r}
sales[,3,drop=F] # column indexing. drop=F makes it return it as a column
```
```{r}
sales[,3]
```

```{r}
sales[4]
```
```{r}
j=sales[c(1,3),]
j
```
```{r}
my_sales=j
my_sales
```
```{r}
colnames(sales)=c('2019','2020','2012')
rownames(sales)=c('tiffany','funlola','yemi')
sales
```
```{r}
sales['tiffany',]
```
```{r}
sales[,'2020']
```
```{r}
sales[,'2020',drop=F]
```
#### matrix Arithmetic###
```{r}
shola <- matrix(c(66.5, 56.0, 75.0, 65.4, 67.9, 70.9, 81.2, 50.0, 76.3 ),
        nrow = 3, byrow = T,
        dimnames = list(c("tiffany", "funlola",  "fatimat"),
                        c("2019", "2020", "2021")))
shola
```
```{r}
scaled_shola=shola/100
scaled_shola
```
```{r}
height <- matrix(c(1.3,6.4,4.5,1.6,2.4,3.3,4.6,1.2,3.0),nrow=3)
height
```
```{r}
bmi=shola/(height)^2
bmi
```
```{r}
colSums(shola) #its a function used to sum up all the entries of a column
```
```{r}
rowSums(shola)
```
```{r}
colMeans(shola) # used to calculate the average weight of all the entries of the column i.e in each year
```
```{r}
Total=colSums(shola)
Average=colMeans(shola)
metrics=rbind(shola,Total,Average)
metrics
```
```{r}
A=matrix(c(13,20,27,51,45,71,19,60,70),nrow=3)
A
``` 
```{r}
A[c(2,3),1,drop=F]
```
```{r}
A[c(2,3),1]
```
```{r}
A[c(2,3),c(2,3)]
```
```{r}
A[c(1,3),c(1,2)]
```
```{r}
A[5]
```
###### Relational and Logical Operations in R#####
```{r}
12>2
```
```{r}
2>20
```
```{r}
print(3!=3)
```
```{r}
print(3==3)
```
```{r}
TRUE==TRUE
FALSE==TRUE
TRUE!=FALSE
```
#### The 'and' operator returns false when either of the values is false and the 'or' operator returns true when either of the values is true.
```{r}
w=9
w<12
w>6
(w<12) & (w>6)
```
##### using the 'not' operator. not operator is !
```{r}
!TRUE
!(4<3)
```
#### If,Else and Elif statement
```{r}
number=-3
if(number<0){
  print('Your number is Negative')}
```
```{r}
v=-1
if(v<0){
  v=v*-1
  print(v)}       
```
```{r}
v=-1
if(v<0){
  v=v*-1
ifelse(v>0,'Yes','No')}
```
```{r}
ifelse(1==1,'Yes','No')
ifelse(1==0,'Yes','No')
ifelse(3>=1,'True','False')
```
```{r}
totest=c(1,1,0,1,0,1)
ifelse(totest==1,'Yes','No')
ifelse(totest==1,totest*3,totest)
ifelse(totest==1,totest*3,'zero')
```
```{r}
v=-7
if(v<0){
  v <- v*-1
  print('i have made your object positive,hey,look!')
  print(v)
} else{
  print('your object is already positive')}
```
```{r}
v <- -7
if(v<0){
  print('The object was less than o,but i am working in it,')
  print('wait for it......')
  v <- v*-1
  print('Your object is now positive,check it out!')
  print(v)
} else if(v==0){
  print('Your object is exactly zero.')
} else if((v>0)&(v<12)){
  print('The object is positive and less than 12.')
} else{
  print('Your object is positive and larger than 12.')
  print(v)
}

```
```{r}
z=c(1,0,-3,5)
if (z>3){
  print('Dooshima has a good handle of R,')
} else{
  print('she is still a learner.')
}
```
```{r}
z <- 'jamal'
if(z=='patricia'){
  print('patricia is a DataLab intern;let her in.')
} else if (z!='patricia'){
  print('i do not know this person')
  }
```

```{r}
temp <- 20
if(temp<=0){
  print('The weather is freezing!')
} else if(temp<=10){
  print('The weather is cold!')
} else if(temp<=20){
  print('The weather is cool')
} else if(temp<=30){
  print('The weather is warm')
} else {
  print('It is hot,mehn!')
  }
```
###### Taking input from user in R###
taking input using 'readline()'.this command will prompt you to input a desired value
```{r}
age <- readline(prompt='Enter your age,please: ');
age <- as.integer(age);
print(paste('Your age is',age))

```
##### functions in R
```{r}
say.hello <- function(){
  print('Hello World')
}
say.hello()
```
```{r}
power=function(x,y){
  result <- x^y
  print(paste(x,'raised to the power', y, 'is',result))}
power(10,3 )
```
 ###### Task######
```{r}
HCF=function(a,b){
  if(a>b){
    lesser=b
  }else{
    lesser=a}
  for(i in 1:lesser){
     if((a%%i==0) & (b %%i==0)){
        HCF=i
     }
  }      
  print(paste('The HCF of',a, 'and',b, 'is', HCF))}
HCF(60,72)
  
```
```{r}
perimeter=function(l,b){
  result=2*(l+b)
  print(paste('The perimeter of the rectangle ABCD is', result))}
perimeter(15,7)
```



```{r}

```
```{r}

```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


i```{r}
```





